\section{Additional information}

\subsubsection*{Code and data availability}

The code will be at
    \href{https://github.com/HAN-CELL-LAB/PhamHansel-intrinsic-exc-2022}{\textit{github.com/HAN-CELL-LAB/PhamHansel-intrinsic-exc-2022}}.
The \textit{in vitro} S1 data have been collected in the Hansel laboratory \citep{Gill2020-wy}
    and will be made available online at the same URL address.
The \textit{in vivo} V1 data are obtained from Dr. Nicholas J. Priebe \citep{Li2020-ej}.
\NEWCHANGE{All experimental data and all essential modeling data are included in the manuscript figures}.

\subsubsection*{Competing interests}

\NEWCHANGE{The authors have no conflicts of interest to declare}.

\subsubsection*{Author contributions}

\NEWCHANGE{The work was performed in the Hansel laboratory at the University of Chicago.
TP contributed to the design of the work,
    the acquisition and analysis of the data and to drafting the work.
CH contributed to the conception and design of the work,
    the interpretation of the data and to the drafting of the work.
Both authors approved the final version of the manuscript,
    and agree to be accountable for all aspects of the work.
All persons designated as authors qualify for authorship,
    and all those who qualify for authorship are listed}.

\subsubsection*{Acknowledgement}

We thank Dr. Nicholas J. Priebe (University of Texas, Austin)
    for providing us with the in vivo V1 data \citep{Li2020-ej}.
We thank Drs. Yali Amit, Brent Doiron, and Stephanie Palmer (University of Chicago)
    for their feedback on the manuscript.
\NEWCHANGE{We are grateful to Silas E. Busch for help with patch-clamp recordings from mouse S1 cortex}.
