\subsection{Recognition of known / learned patterns in a simple feedforward two-layer binary network}

\input{figures/Fig3}

To extend the framework, we set up our model as a simple feedforward network
    with an input layer $X$ and an output layer $Y$,
    connected via the non-negative connectivity matrix $W_{YX}$
        that describes the amount of selectivity of the inputs to each output unit via
            the selectivity index $W$ (selectivity of the output unit $Y$)
            and a specified amount of input overlap (overlap in pattern presentation in the input unit $X$)
            (\autoref{eq:ffwdweight-alphaW}, \autoref{fig:demo-ffwd-recog}a,b).

The (in)completeness of the inputs is characterized by
        the \% complete parameter, with added noise.
    The output units at $Y$ are activated with a Heaviside step function (\autoref{eq:heavisidefun}), whose thresholds are characterized by
        the base thresholds $\theta_{\text{base}}$
        and an additive threshold change $\Delta \theta$ in only one of the output units (\autoref{fig:demo-ffwd-recog}c).
    $\Delta \theta < 0$ would be analogous to an increased intrinsic excitability of the affected artificial neuron compared to the other output units,
        whereas $\Delta \theta > 0$ would indicate decreased intrinsic excitability.
    Here, we look at the effects of $\Delta \theta$ on the performance in a recognition task
        (focus on the output whose threshold is changed) (\autoref{fig:demo-ffwd-recog}c).

\input{figures/Fig4}

The results are shown in \autoref{fig:ffwd-recog-perf}a,
        recognition true positive rate (TPR) is shown in blue,
        while the tradeoff between TPR and the false-positive rate, FPR, (TPR $-$ FPR) is shown in red.
    Because $N_Y=3$, $\alpha_W < 1/3$ would be lower than chance,
        and thus highly undesirable for recognition.
    Increased selectivity emphasizes relevant inputs and reduces the weight of irrelevant ones, thus benefitting both performance parameters.

The dependence of recognition performance on $\Delta\theta$ (true positive rate; TPR)
    appears sigmoidal, inverted with the sign of threshold change
        (blue shaded lines in \autoref{fig:ffwd-recog-perf}a).
    As expected, lowered thresholds allow for easier recognition of the output of interest, eventually reaching 100\% TPR.
    When the selectivity index is high and the inputs are mostly complete, the unit does not require decreased thresholds to reach maximal recognition.
    However, for a given input completeness, weaker selectivity requires a more pronounced threshold decrease to reach a set performance level.
    The same is true when inputs become more incomplete, given a weak to moderate $\alpha_W$.
    Hence the optimal threshold change for maximal recognition leans toward stronger intrinsic potentiation $\Delta \theta < 0$
        when either the inputs become more insufficient or the network selectivity strength $\alpha_W$ is imperfect (\autoref{fig:ffwd-recog-perf}
        \NEWCHANGE{c,d}).

\NEWCHANGE{Generally, FPR monotonically increases with decreases in threshold values.
    Overall, decreased input completeness leads to decreases of FPR
        as irrelevant inputs will have less of a chance
        to evoke responses in the relevant output unit $Y$.
    This would be more beneficial for recognition performance
        (see ROC curves in {\autoref{fig:ffwd-recog-perf}}b).
}

As the goal is to maximize recognition performance without enhancing the false recognition rate,
        we want to maximize TPR while minimizing FPR,
        thus maximizing the difference TPR $-$ FPR
        (red shaded lines in \autoref{fig:ffwd-recog-perf}a).
    When selectivity is too low
        ($ < 1/3$, leftmost top sub-panel in \autoref{fig:ffwd-recog-perf}a
        and lightest grey line in the top right sub-panel in
        \autoref{fig:ffwd-recog-perf}\NEWCHANGE{d}$_2$),
        FPR is higher than TPR as irrelevant input strengths are now stronger than relevant connections,
        hence TPR $-$ FPR $<$ 0.
    In moderate selectivity strengths, these difference curves
        are negatively affected by increased input incompleteness
        and require more negative $\Delta \theta_{\mathrm{optim}}$ to reach their maximum
        (\autoref{fig:ffwd-recog-perf}\NEWCHANGE{d}).

\subsubsection*{Effects of input selectivity overlap}

Aside from the imperfection of selectivity ($\alpha_W < 1$),
        another possible issue that could arise intrinsically in the feedforward setup is
        an overlap in input selectivity, as certain input units could activate more than one output.
    Here, we consider the worst case, in which some units connect strongly to all outputs (\autoref{fig:demo-ffwd-recog}b).
    As we assume a fixed selective input number per output ($N_{\text{selec}} = 50$),
        nonzero overlap input units send relevant,
        but also irrelevant input to the output units.
    When selectivity strength is low, this reduces the relative contribution of relevant input.

As defined, recognition TPR is not really affected by the selectivity overlap,
        as we focus only on output units encoding the learned object pattern
        (\autoref{fig:ffwd-recog-perf}a).
    This is because the high overlap allows the shared inputs
        to evoke responses in the affected output too easily
        when the input pattern actually belongs to another output class.
    For this reason, sparsity of input patterns is more beneficial
        for the tradeoff between TPR and FPR (i.e. the difference) when $\Delta \theta < 0$.
    On the other hand, when $\Delta \theta > 0$, input completeness is better for the recognition tradeoff.
    However, if we look at the optimal threshold change,
        as input patterns become more incomplete and have a moderate selectivity index,
        $\Delta \theta_{\mathrm{optim}}$ generally tends towards more negative changes
        in the case of nonzero selectivity overlap
        (\autoref{fig:ffwd-recog-perf}\NEWCHANGE{d}, bottom sub-panel).

A well-known analysis tool for examining binary classification performance is the ROC graph.
    As expected,
        increased input completeness
        would increase recognition performance
        \NEWCHANGE{{\autoref{fig:ffwd-recog-perf}b}}.
    Increase in selectivity overlap decreases the overall performance of the recognition task
        (compare top and bottom panels of
        \NEWCHANGE{{\autoref{fig:ffwd-recog-perf}b}}).

\subsection{Biological threshold potentials and activity-dependent plasticity}

The purpose of re-analyzing existing biological data sets is twofold.
    First, we would like to obtain information on the distribution of threshold values
        and the difference between resting and threshold potentials in cortex.
    Second, as our modeling has shown, negative changes in threshold potentials
        can signify a shift in neural encoding strategies
        from accurate reporting of input patterns
        to interpretative reporting of whether a known / previously learned pattern is presented.
    In this latter scenario, it is of interest,
        which biological activity pattern drives this threshold adaptation
        and thus enables optimal signal processing
        when the history of the input pattern presentation (previously presented under favorable conditions or not) changes.

\subsubsection*{Threshold potentials obtained from V1 \textit{in vivo} recordings}

\input{figures/Fig5}

We obtained resting and threshold potentials
        from published whole-cell patch-clamp recordings
        from L2/3 pyramidal neurons in the primary visual cortex (V1)
        of awake macaque monkeys \citep{Li2020-ej}.
    These recordings were performed
        in the laboratory of Dr. Nicholas Priebe (University of Texas at Austin)
        and requested measures have been made available to us.

In this data set, resting potentials $V_R$ ranged from about $-75$ mV to $-50$ mV
            ($-64.83 \pm 6.39$ mV,
            $n = 26$ cells from 3 animals)
        and threshold potentials $V_T$ ranged from about $-60$ mV to $-40$ mV
            ($-51.48 \pm 4.69$ mV,
            $n = 26$).
    The difference between resting and threshold potentials
        $\Delta V_{TR}$ in these recordings (the distance in mV)
            \NEWCHANGE{ranged from $5$ to $22$ mV}
            ($13.34 \pm 4.48$ mV,
            $n = 26$)
        (\autoref{fig:ffwd-biol-data}).
    This analysis shows the range of threshold potentials assumed
        in a biological setting and under physiological conditions
        in the cortex of the awake macaque,
        an animal model that is as close as it gets to the human cortex.

\subsubsection*{Threshold potentials and plasticity in S1 \textit{in vitro} recordings}

Our own \textit{in vitro} recordings
        from the primary sensory cortex (S1) of the mouse \citep{Gill2020-wy}
        allow us to perform more detailed analyses,
        including the analysis of activity parameters that
        alter the threshold potential.
    The data are shown in \autoref{fig:ffwd-biol-data}b.
    We observe that the resting potentials $V_R$ in S1 assume values
        that are similar to those seen in the V1 cortex of awake macaques,
        roughly in the range of $-75$ mV to \NEWCHANGE{$-45$} mV
            \NEWCHANGE{(pre $= -65.50 \pm 5.25$ mV,
                post $= -65.73 \pm 6.72$ mV,
                paired pre-vs-post $t$-stat $= 0.3807$,
                $p = 0.7063$,
                $n = 29$ cells)}.
        \NEWCHANGE{We did not observe significant differences between
            the resting potentials of S1 (combined pre \& post) and V1 data
                ($t$-stat $= -0.5433$,
                $p = 0.5884$)}.
    However, the threshold potentials $V_T$ occupy
        a more positive range than those observed in the V1 recordings;
        we observed a range from about $-50$ mV to \NEWCHANGE{$-30$} mV
            \NEWCHANGE{(pre $= -42.14 \pm 2.59$ mV,
                post $= -42.22 \pm 4.00 $ mV,
                paired $t$-stat $= 0.1311$,
                $p = 0.8966$,
                $n = 29$),
            as well as significant difference between S1 and V1 threshold potentials $V_T$
                ($t$-stat $= 10.3571$,
                $p = 1.4765 \times 10^{-16}$)
            },
        leading to a larger difference between resting and threshold potentials
            \NEWCHANGE{($\Delta V_{TR}$) in S1 than in V1
                (V1 $= 13.34 \pm 4.48$ mV,
                S1 pre $= 23.36 \pm 5.70$ mV,
                S1 post $= 23.51 \pm 7.80$ mV;
                V1-vs-S1 two-sample $t$-stat $= 6.9342$,
                $p = 8.6032 \times 10^{-10}$)}.
    \NEWCHANGE{A caveat of our V1 to S1 comparisons certainly is that
            the values were taken from different recording conditions
            (\textit{in vivo} for V1,
            \textit{in vitro} -- without an attempt to ``activate'' the slice -- for S1).
        Note, however, that the}
        finding of a larger $\Delta V_{TR}$ corresponds
        to observations made in S1 cortex of intact rats \citep{Brecht2003-vf}.

The \textit{in vivo} V1 data set \citep{Li2020-ej} from the Priebe laboratory does not contain plasticity data,
        but our own published recordings from L2/3 pyramidal neurons in S1 mouse cortex \textit{in vitro} do \citep{Gill2020-wy}.
    As reported in \cite{Gill2020-wy},
        three manipulations lead to increases in the measure of excitability used in this work,
        the number of spikes evoked by somatic injection of a depolarizing current test pulse.
    In a total of
        \NEWCHANGE{29  recordings (23 from that study and 6 additional recordings)},
        the experimental manipulations and number of recorded cells per manipulation are as follows:
        electric activation (repeated injection of depolarizing currents or tetanic activation of excitatory synapses; $n=10$);
        bath application of the muscarinic agonist oxo-m ($7 \mu M$; \NEWCHANGE{$n=10$});
        paired oxo-m application with electric stimulation (\NEWCHANGE{$n=9$}).
    All three protocols cause an increase in excitability \citep{Gill2020-wy}
        and therefore qualify as efficient IP protocols \autoref{fig:ffwd-biol-data}d.
    Changes in threshold potential have not been analyzed and reported in \cite{Gill2020-wy},
        but are examined here.

\NEWCHANGE{Significant negative threshold changes $V_T$ were observed
    with paired electric and cholinergic activation
        ($-1.96 \pm 2.37$ mV,
        paired Student's $t$-stat $= 2.4897$,
        $p = 0.0375$,
        $n=9$),
    but not with electric
        ($0.23 \pm 2.28$ mV,
        $t$-stat $= -0.3210$,
        $p = 0.7555$,
        $n = 10$)
    or cholinergic activation alone
        ($1.29 \pm 4.76$ mV,
        $t$-stat $= -0.8549$,
        $p = 0.4148$,
        $n = 10$).
    Thus, cholinergic co-activation provides the most efficient means
        to trigger a lowering of the spike threshold.
}

We cannot fit the thresholds observed directly to our model shown in \autoref{fig:ffwd-recog-perf}a,
        but made an attempt to approximate minimal and maximal threshold potential values
        that correspond to the $-0.5$ and $0.5$ edges of the model plots.
    As the spike threshold $V_T$ cannot be lower than the resting potential $V_R$,
        we accepted the average resting potential measured in our recordings
        as an approximation of
        $\theta_{\mathrm{min}} = 0, \Delta \theta_{\mathrm{min}} = -0.5$.
    The value obtained from our recordings
        \NEWCHANGE{($n=29$) is $-65.61$ mV}
        (combined pre and post measurements).

As the theoretically available most positive threshold bound
        $\theta_{\mathrm{max}} = 1, \Delta \theta_{\mathrm{max}} = 0.5$,
        we accepted the membrane potential at which action potentials are evoked
            from a depolarizing pulse when the neuron is in a (relative) refractory state.
    In a set of new recordings, this value was determined to be $-30.71$ mV
        ($n=4$; \autoref{fig:ffwd-biol-data}c).
    Therefore, the threshold shifts available for optimization of performance in the recognition task
        are limited to a maximal range of about $35$ mV.
    Note that in our set of
        \NEWCHANGE{$29$ recordings},
        the range of ``real'' (measured in the non-refractory state) threshold potentials ($V_T$) was about $10-15$ mV wide.
    Although this observation does not exclude the possibility that larger changes can take place,
        in particular with repeated instead of single activation periods,
        it seems likely (and plausible) that under physiological conditions
            the range of assumed threshold potentials
            is smaller than the theoretically available one.

\NEWCHANGE{How does this threshold potential influence the spike output in our recordings?}
    First, we notice that threshold plasticity ($V_T^{\Delta}, \Delta V_{TR}^{\Delta}$) is bidirectional,
        even though the change in spike number ($n_{spk}^{\Delta}$) is largely unidirectional.
    This means that in some cells the threshold for spike firing is enhanced,
        but once passed, more spikes can be evoked.
    Second, this bidirectional change leads to a widening of the range of threshold potentials
        after IP is triggered (\autoref{fig:ffwd-biol-data}b).
    \NEWCHANGE{Additionally, we observe significant negative association between
            the changes in spike number ($n_{spk}^{\Delta}$) and
            threshold-to-rest distance changes ($\Delta V_{TR}^{\Delta}$)
            ($\rho = -0.41, R^2 = 0.17, p = 0.0275$),
            but not with the changes of the other two membrane potential variables
            ($ p = 0.0996 $ for $V_{T}^{\Delta}$
            and $ p = 0.3547$ for $V_{R}^{\Delta}$).
            This observation does not imply that $V_T$ and $V_R$ have no impact on spike firing.
            What it shows is that
                -- as can be predicted from the definition of $\Delta V_{TR}$ as
                the actual membrane potential difference that has to be overcome to reach threshold
                -- spike firing is most sensitive to changes in $\Delta V_{TR}$.
    }

Our model on the relationship
        between performance in a recognition task and changes in the threshold potential
        suggests that under a wide range of conditions
        task performance benefits from a reduction in the
        \NEWCHANGE{spike threshold distance, $\Delta V_{TR}$}.
    The experimental data presented in \cite{Gill2020-wy} allow us to examine
        what type of neuronal activation
        causes a negative \NEWCHANGE{$\Delta V_{TR}$} shift
        and therefore alter the neuron's coding strategy to one that favors the recognition of known / learned input patterns,
        in particular when the presented pattern is incomplete.
    To assess differences in the efficacy to promote negative threshold changes,
        we applied $\chi^2$ test of independence
        for the distribution of threshold potential changes
        triggered by the three types of manipulations used (electric, cholinergic and cholinergic paired with electric),
        along two categories: negative and non-negative threshold changes.
    Here, the category ``negative threshold change'' includes
        \NEWCHANGE{all threshold distance changes, i.e. $\Delta V_{TR}^{\Delta}$}
        in a negative direction regardless of amplitude.
    We defined two exceptions:
        a) the negative threshold change category does not include recordings,
            in which there was a decrease in spike number
                that might offset some effects of the threshold change,
        and b) cells with a ``null'' direction change in threshold potential
            were counted as ``non-negative''.
    \NEWCHANGE{Based on the distribution of threshold changes,
        we identified four recordings,
            in which these potential values clustered around zero change,
            and whose percentual change remained $< 1\%$.
        We therefore defined a negative threshold change as
            one exceeding the $1\%$ threshold distance}.
    With this definition of categories,
        in \NEWCHANGE{$7/9$ recordings}
            with paired cholinergic and electric activation
            a negative threshold change was observed,
        while this was only the case in
            \NEWCHANGE{$3/10$ recordings}
            in which cholinergic activation was applied alone,
        and in $1/10$ recordings in which electric activation was applied alone
        (\autoref{fig:ffwd-biol-data}e).
    The paired cholinergic activation was the only activation pattern
        to promote a \NEWCHANGE{significant} negative threshold change
        \NEWCHANGE{($\chi^2 = 9.6504; p = 0.00802$)}.
